---
title: "Hobos"
author: "GMB"
date: "January 23, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## North - South temperarutre

This file was made to explore and analyse the temperature data from Brad Tucker's tires with Aedes spp. eggs. 

```{r}
### Clean workspace
rm(list = ls())

### Load libraries
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)

### Load data
setwd("~/Documents/CBS PhD/albopictus OW/Data")
OW <- read.csv("OWALL_030520.csv")
OWA <- read.csv("OWALL_A030520.csv")
OWAVG <- read.csv("OWALL_AVG030520.csv")

```

The data is loaded and contains `r nrow(OW)` rows and `r ncol(OW)` columns.

Now we manipulate the data so we can look at it over time. 

```{r, fig.height=9, fig.width=9}
OW$Date <- as.Date(OW$Date, "%d-%b-%y")
OW <- subset(OW, Date > "2018-11-18" & Date < "2019-04-10")
OW$MilTime <- format(strptime(paste(OW$Time, OW$AMPM, sep = " "), "%I:%M %p"),"%H:%M")
#OW$MilTime > 12:00 & OW$MilTime <12:30 
#OW$BlockTime <- ifelse(OW$MilTime >0:00 & OW$MilTime < 3:01, "0-3", 
#                ifelse(OW$MilTime >3:00 & OW$MilTime < 6:01, "4-6",
#                ifelse(OW$MilTime >6:00 & OW$MilTime < 9:01, "7-9", 
#                ifelse(OW$MilTime >9:00 & OW$MilTime < 12:01, "10-12", 
#                ifelse(OW$MilTime >12:00 & OW$MilTime < 15:01, "12-15",
#                ifelse(OW$MilTime >15:00 & OW$MilTime < 18:01, "16-18",
#                ifelse(OW$MilTime >18:00 & OW$MilTime < 21:01, "19-21",
#                ifelse(OW$MilTime >21:00 & OW$MilTime < 23:59, "22-24", #"Stinks"))))))))
OW$GMTH <- as.POSIXct(paste(OW$Date, format(strptime(paste(OW$Time, OW$AMPM, sep = " "), "%I:%M %p"),"%H"), ":00", sep = ""), tz="GMT")
OW$GMT <- as.POSIXct(paste(OW$Date, OW$MilTime), tz="GMT")
kable(table(OW$Location, OW$Date), caption = "Number of observatinos by Site and Date")

OW$ABC <- ifelse(OW$Number == 1 |OW$Number==4 | OW$Number==7 | OW$Number==10 | OW$Number==13, "A", 
           ifelse(OW$Number == 2 |OW$Number==5 | OW$Number==8 | OW$Number==11 | OW$Number==14, "B", "C") )

g <- ggplot(data = OW, aes(x = GMT, y = Value))
g + geom_line(aes(linetype = Type, color = Type)) + 
  geom_hline(yintercept=-12, linetype = 2)+
  facet_grid(Location~ABC) + theme_classic()+
  scale_colour_manual(values=c("black","red"))+
  ggtitle("Continous temperature reads by tire location")
```

We reduced the data set from `r min(OW$Date)` to `r max(OW$Date)`.
And renamed number 1 to 15 with ABC

A = 1, 4, 7, 10, 13
B = 2, 5, 8, 11, 14
C = 3, 6, 9, 12, 15

Next we want to answer some of the questions:

1. What is the temperature difference between Ambient and Tire each day
2. How many observation were below -12C?
3. How many freeze - thaw cycles occur
4. ....

I can't find my notes from our previous meeting with the exact questions. 


```{r}
### Ambient v. Tire
ByHour <- summarize(group_by(OW, Location,ABC, GMTH, Type), N = n(), C = mean(Value))
OWmelt <- dcast(ByHour, Location+ABC + GMTH ~ Type)
OWmelt <- subset(OWmelt, !is.na(Ambient)) # lose ~50%
OWmelt <- subset(OWmelt, !is.na(Tire)) # lose another 75%
OWmelt$Dif <- OWmelt$Ambient - OWmelt$Tire

h <- ggplot(data = OWmelt, aes(x = GMTH, y = Dif))
h + geom_line(aes()) + 
  geom_hline(yintercept=0, linetype =2)+
  facet_grid(Location ~ ABC) + theme_classic()+
  ggtitle("Difference between ambient and tire temp by hour (negative = tire is warmer)")
```
We tried to summarize by hour, this did not work very well. We lost a lot of observations.  Next we tried by day. 

```{r, fig.height=9, fig.width=9}
ByDay <- summarize(group_by(OW, Location,ABC, Date, Type), N = n(), C = mean(Value), minT = min(Value), maxT = max(Value))
OWmelt_D <- dcast(ByDay, Location+ABC + Date ~ Type)
OWmelt_D <- subset(OWmelt_D, !is.na(Ambient)) 
OWmelt_D <- subset(OWmelt_D, !is.na(Tire)) 
OWmelt_D$Dif <- OWmelt_D$Ambient - OWmelt_D$Tire

k <- ggplot(data = OWmelt_D, aes(x = Date, y = Dif))
k + geom_line(aes()) + 
  geom_hline(yintercept=0)+
  facet_grid(Location~ABC) + theme_classic()+
  ggtitle("Difference between ambient and tire by day (negative = tire is warmer)")

```
```{r, fig.height=9, fig.width=7}
l <- ggplot(data = ByDay, aes(x = Date, y = minT))
l + geom_line() + 
  geom_hline(yintercept=-12, linetype =2)+
  facet_grid(Location + ABC ~ Type) + theme_classic()+
  ggtitle("Minumum temperature per day")

```


```{r, below 12}
#### Below - 12
OW$Min10 <- ifelse(OW$Value < -12, 1, 0)
PercMin10 <- summarize(group_by(OW, Location, ABC, Type), Min10 = sum(Min10), N = n(), Perc = round(sum(Min10)/n()*100, 2))
PercMin10_cast <- dcast(data = PercMin10, Location + ABC ~ Type)
kable(PercMin10_cast, caption = "Percentage of reads below -12C")

subset(OW, Location == "New Berlin" & ABC == "B" & Type  == "Tire" & Value < -12)
subset(OW, Location == "NWMAD" & ABC == "C" & Type  == "Tire" & Value < -12)

```

In New Berlin B, there were 3 reads below 12C
In NWMAD C there was 1 read below 12C. 


It does hit -13.5 and -14. 


## Tables

```{r}
library(lubridate)
OW$week <- week(OW$GMT)
OW$year <- year(OW$GMT)
OW$month <- month(OW$GMT)

Tab1 <- summarize(group_by(OW, Location, Type, year, month), minT = min(Value),
          maxT = max(Value), avgT = round(mean(Value),1), sdT = round(sd(Value),1))
Tab1$Temps <- paste(Tab1$avgT, " (", Tab1$minT, ", ", Tab1$maxT, ")")
Tab1_melt <- dcast(Tab1, Location+Type~ year+month, value.var = "Temps")
kable(Tab1_melt, caption = "Average (min and max) temperatur per month")

#Update

# TAble 2
OWmelt_D$week <- week(OWmelt_D$Date)
OWmelt_D$year <- year(OWmelt_D$Date)
OWmelt_D$month <- month(OWmelt_D$Date)
Tab2 <- summarize(group_by(OWmelt_D, Location, ABC, year, month), minT = min(Dif),
          maxT = max(Dif), avgT = round(mean(Dif),1), sdT = round(sd(Dif),1))
Tab2$Diff <- paste(Tab2$avgT, " (", Tab2$minT, ", ", Tab2$maxT, ")")
Tab2_melt <- dcast(Tab2, Location+ABC~ year+month, value.var = "Diff")
kable(Tab2_melt, caption = "Average (min and max) difference per month")
```

##IGNORE

Wroking with Time Series package in R you can pull all kinds of information. Lets see if it works. 

```{r}
#plot.ts(OW$Value)
#plot.ts(log(OW$Value))

#library(TTR)
#Test <- SMA(OW$Value, n=24)
#plot.ts(Test)
#plot(as.ts(OW$Value)) 

#Set1 <- subset(OW, Location == "SCCMAD" & Number == 4 & Type == "Tire")
#Set1 <- subset(Set1, !is.na(Value))
#plot(SMA(Set1$Value))


#### Exploring "trends"

## Old sckool
# https://anomaly.io/seasonal-trend-decomposition-in-r/
#library(forecast)
#trend_beer = ma(Set1$Value, order = 8, centre = T)

#plot.ts(Set1$Value)
#lines(trend_beer)

#detrend_beer = Set1$Value / trend_beer
#plot.ts(detrend_beer)

## using decompose
#ts <- ts(Set1$Value, frequency = 8*30)
#ts_decom <- decompose(ts, "multiplicative")
#plot(ts_decom)


#### Autocorrelation
#acf(Set1$Value) # lag of 8 has max correlation
#Set1DF <- diff(Set1$Value, differences = 2)
#plot(Set1DF)
#Set1DFDF <- diff(Set1DF, lag = 8)
#plot(Set1DFDF)
```


All nice and well. Lets get back to the questions. 

There are 3 tire stacks at each location. For some reason there aren't matching data for each time frame and tire stack, we lost 12000 observations when trying to match. 

```{r}

```